### 2022-07-10 chapters/benchmarking/draft.tex
correction : chapter 3 part 1/2

\section{Motivation}
Empirical measurements are critical to capture the effect of developers' choices on software energy consumption.
To accomplish this, one should not overlook the benchmarking pitfalls highlighted in the state of the art.
Furthermore, when it comes to evaluating the energy consumption of software, other challenges occur, such as the impact of hardware, the difference between the machines used for measurements and those that may apply the results, and so on.
Moreover, considerable advances in computer science have resulted in an increase in the number of obsolete results.
Addressing all of them becomes significantly more difficult in studies that compare multiple treatments.

Furthermore, between the exploratory experiment and the published results, there may have been the appearance of new candidates as well as the evolution of others.
As a result, it is not only necessary to assure the "reproducibility" of the results, but it is also essential to provide room for new candidates.
As a result, we'd like to present "Extension," a new approach for leveraging experimentation.
The ability to give not only the tools needed to replicate the tests but also to integrate more candidates, workloads, or important key performance indicators

This chapter covers ways to overcome empirical analysis challenges in energy consumption studies. First, we'll go through the three components of a successful benchmark within the energy consumption experiment. Section~\ref{sec:benchmarking_reproducibility} will therefore concentrate on the "reproducibility" issue and how to deliver reproducible experiments without interfering with the energy measurement. Section~\ref{sec:taming-the-energy-variation} will discuss accuracy in software energy consumption. Finally, Section~\ref{sec:extension} will discuss tools for introducing new candidates, workloads, or other key performance indicators.

@@

This option, however, comes at a cost.

Because of the hypervisor, software will be built on two kernels: one for the virtual machine (guest) and one for the host machine, resulting in a visible overhead and a negative influence on the performance of the system-under-test.

As a result, we cannot use VM for performance-related tests.

Isolation is another drawback of VM.

While this feature protects the experimental setting from unwanted interference from the outside world, it is possible that this interaction is required—-especially if the experiment is dependent on an external source such as energy montitors .

@@

Containers provide such an advantage by ensuring application separation and ease of replication.
Figure~\ref{environement:virtualization_technique}depicts the architectural differences between virtualization and container technologies. There are three main types of virtualisation.
\begin{itemize}
    \item \textsf{Type\,1}: runs on the hardware directly.
          It is primarily utilized by cloud providers where there is no host OS and only VMs that run on the open-source Xen or VMware ESX hypervisors.

    \item \textsf{Type\,2}: runs on top of the host operating system and is most commonly found on personal computers.

          VMware server and virtualBox are notable examples of this category, and most researchers' experiments use them.
          However, because of the two operating systems, the applications are typically slower.

    \item \textsf{Containers}: run their operating systems on the host kernel rather than their own , which makes them smaller, faster, and more efficient in terms of hardware utilization.
          one can cite \emph{Docker}, \emph{Linux LXC}, or \emph{LXD}~\cite{abuabdo_virtualization_2019}.
\end{itemize}

@@

\subsection{Docker Vs. Virtual Machine}
Despite the fact that Type 1 is more performant than Type 2, the latter is the most used in research, as most researchers tend to conduct their experiments on their own machines. Docker, on the other hand, is the most well-known container technology. In our case, we are more likely to promote Docke for two reasons:
\begin{enumerate}
    \item As previously stated [143] [reference Morabito (2015) and van Kessel et al. (2016)], we require a lightweight orchestrator to limit the overhead on energy usage of our experiments
          \cite{van2016power}
          % TODO : ask romain about the title of the paper and the reference to the paper
          [cite Morabito (2015) and van Kessel et al. (2016)],
          % cite Power efficiency of hypervisor-based virtuali- zation versus container-based virtualization. University of Amsterdam.
    \item We need to communicate with the host OS because we are using hardware sensors to measure the energy consumption..
\end{enumerate}

A special notice to \href{https://github.com/powerapi-ng/virtualwatts}{virtualwatts}. A framework that enables us to retrieve the energy consumption of a virtual machine.

@@

Using research from \cite{santos2018does} who examined how adding the Docker layer affected energy consumption.
\citeauthor{santos2018does}~\emph{et~ al.}conducted their experiment by running numerous benchmarks both with and without Ddocker.They contrasted the energy usage and execution time that resulted.
The first step was to observe the effects of the orchestrator and the Docker deamon while there was no work to be done.
Then, they use the following benchmarks in their experiments:
%%

\begin{itemize}
    \item Wordpress,
    \item Reddis,
    \item PostgreSQL.
\end{itemize}

The values below show the system under test's energy consumption while it is idle.
Docker has an overhead of roughly $1,000$  joules, as seen in Figure~\ref{fig:docker_idle}

@@

To summarize, Docker-based software tendst to consume more energy since it takes longer to execute.
The execution of the Docker deamon causes an increase in average power consumption of only \textbf{2\,Watts}.
This overhead can reach up to 5\% in IO-intensive applications, while it is barely visible in CPU- or DRAM-intensive workloads.

@@

\subsection{Conclusion}
As can be seen, the introduction of the supervisor has increased the impact of Docker on energy.
which will be applied equitably to all experiments.
Therefore when it comes to comparison analysis, it will mitigate its impact autimatically.
Furthermore, because we have access to the host Hardware, we do not need to worry about capturing the SUT's energy use.
As a result, we will use Docker as an encapsuation solution for all of our experiments.
It would ensure the reproducibility of our experiments in a simple and straightforward manner.
\clearpage

>>>>

\section{Motivation}
Empirical measurements are critical to capture the effect of developers' choices on software energy consumption.
To accomplish this, one should not overlook the benchmarking pitfalls highlighted in the state of the art~\cite{van_der_kouwe_benchmarking_2018}.
Furthermore, when it comes to evaluating the energy consumption of software, other challenges occur, such as the impact of hardware, the difference between the machines used for measurements and those that may apply the results, and so on.
Moreover, considerable advances in computer science have resulted in an increase in the number of obsolete results.
Addressing all of them becomes significantly more difficult in studies that compare multiple treatments.

Furthermore, between the exploratory experiment and the published results, there may have been the appearance of new candidates as well as the evolution of others.
As a result, it is not only necessary to ensure the "reproducibility" of the results, but it is also essential to provide room for new candidates.
As a result, we'd like to present "Extension," a new approach for leveraging experimentation.
The ability to give not only the tools needed to replicate the tests but also to integrate more candidates, workloads, or important key performance indicators

This chapter covers ways to overcome empirical analysis challenges in energy consumption studies. First, we  go through the three components of a successful benchmark within the energy consumption experiment. Section~\ref{sec:benchmarking_reproducibility}  therefore concentrates on the "reproducibility" issue and how to deliver reproducible experiments without interfering with the energy measurement. Section~\ref{sec:taming-the-energy-variation}  discusses accuracy in software energy consumption. 
Finally, Section~\ref{sec:bench_extension} presents tools for introducing new candidates, workloads, or other key performance indicators.

@@

This option, however, comes at a cost. Because of the hypervisor, software will be built on two kernels: one for the virtual machine (guest) and one for the host machine, resulting in a visible overhead and a negative influence on the performance of the system-under-test. As a result, we cannot use VM for performance-related tests. Isolation is another drawback of VM. While this feature protects the experimental setting from unwanted interference from the outside world, it is possible that this interaction is required—-especially if the experiment is dependent on an external source such as energy montitors .

@@

Containers provide such an advantage by ensuring application separation and ease of replication.
Figure~\ref{environement:virtualization_technique} depicts the architectural differences between virtualization and container technologies. There are three main types of virtualisation.
\begin{itemize}
    \item \textsf{Type\,1}: runs on the hardware directly.
          It is primarily utilized by cloud providers where there is no host OS and only VMs that run on the open-source Xen or VMware ESX hypervisors.

    \item \textsf{Type\,2}: runs on top of the host operating system and is most commonly found on personal computers. VMware server and virtualBox are notable examples of this category, and most researchers' experiments use them.
          However, because of the two operating systems, the applications are typically slower.

    \item \textsf{Containers}: run their operating systems on the host kernel rather than their own , which makes them smaller, faster, and more efficient in terms of hardware utilization. One can cite \emph{Docker}, \emph{Linux LXC}, or \emph{LXD}~\cite{abuabdo_virtualization_2019}.
\end{itemize}

@@

\subsection{Docker Vs. Virtual Machine}
Despite the fact that Type 1 is more performant than Type 2, the latter is the most used in research, as most researchers tend to conduct their experiments on their own machines. Docker, on the other hand, is the most well-known container technology. In our case, we are more likely to promote Docker for two reasons:
\begin{enumerate}
    \item As previously stated [143] [reference Morabito (2015) and van Kessel et al. (2016)], we require a lightweight orchestrator to limit the overhead on energy usage of our experiments
          \cite{van2016power}
          % TODO : ask romain about the title of the paper and the reference to the paper
          [cite Morabito (2015) and van Kessel et al. (2016)],
          % cite Power efficiency of hypervisor-based virtuali- zation versus container-based virtualization. University of Amsterdam.
    \item We need to communicate with the host OS because we are using hardware sensors to measure the energy consumption.
\end{enumerate}

%  TODO : Rephrase this 
% A special notice to \href{https://github.com/powerapi-ng/virtualwatts}{virtualwatts}. A framework that enables us to retrieve the energy consumption of a virtual machine.

@@

Using research from \cite{santos2018does} who examined how adding the Docker layer affected energy consumption. \citeauthor{santos2018does}~\emph{et~ al} conducted their experiment by running numerous benchmarks both with and without Ddocker. They contrasted the energy usage and execution time that resulted.
The first step was to observe the effects of the orchestrator and the Docker deamon while there was no work to be done.
Then, they use three benchmarks in their experiments: Wordpress, Reddis and PostgreSQL.
The values below show the system under test's energy consumption while it is idle.
Docker has an overhead of roughly $1,000$  joules, as seen in Figure~\ref{fig:docker_idle} .

@@

To summarize, Docker-based software tends to consume more energy since it takes longer to execute.
The execution of the Docker deamon causes an increase in average power consumption of only \textbf{2\,Watts}.
This overhead can reach up to 5\% in IO-intensive applications, while it is barely visible in CPU- or DRAM-intensive workloads.

@@

\subsection{Conclusion}
As can be seen, the introduction of the supervisor has increased the impact of Docker on energy that will be applied equitably to all experiments.
Therefore when it comes to comparison analysis, it will mitigate its impact automatically.
Furthermore, because we have access to the host Hardware, we do not need to worry about capturing the SUT's energy use.
As a result, we will use Docker as an encapsuation solution for all of our experiments.
It will ensure the reproducibility of our experiments in a simple and straightforward manner.
\clearpage

### 2022-07-10 chapters/benchmarking/experiment.tex
correction : chapter 3 part 1/2

\newpage
% To satisfy those three requirements. we elaborated a benchmarking protocol that aims to help reaserchers and practicioners in their experementations.
% for each aspect we present a set of guidelines and tools.
%% This is a conclusion maaan  !! 
\section{Improving accuracy by Taming the energy variation}\label{sec:taming-the-energy-variation}

While the previous section aimed to ensure the reproducibility of our software energy consumption experiments, this section will provide a collection of tips and tools to help you increase the accuracy of these experiments.
We are well aware of the effect that hardware has on energy fluctuations.
However, we feel that there is still an opportunity for practitioners to minimize this energy variance by employing solely tuneable factors.
To that end, we conducted a series of empirical studies utilizing state-of-the-art recommendations to discover which controllable elements can limit the variations of benchmark energy usage.
% We will first start with evaluating the parameters mentioned by \cite{heinrich_predicting}.

In this section we will try to answer the following Research questions.
\begin{description}
    \item[\textsc{RQ}~1:] Does the benchmarking protocol affect the energy variation?
    \item[\textsc{RQ}~2:] How important is the impact of the processor features on the energy variation?
    \item[\textsc{RQ}~3:] What effect does the operating system have on energy variation?
    \item[\textsc{RQ}~4:] Does the choice of processor make a difference in reducing the energy variation?
\end{description}

@@

We used the platform Grid5000 (G5K)~\cite{grid5000,margery:hal-00965708}, which is a large, flexible testbed for experiment-driven research that is spread out across all of France, to meet these needs.Grid5000 has several clusters that are made up of 4 to 124 identical machines with different configurations for each cluster.For our experiment, we looked at four groups. Our main criterion was the CPU configuration.
The table below~\ref{table:g5k}, presents a description of the four cluster that have been chosen for our experiments.

@@

Figure~\ref{fig:topo}  shows the detailed topology of a node in the cluster \textsf{Dahu}. as an example.

@@

First, \textbf{scalability}: We wanted to learn as much as possible about the experiment in the time we had, so we needed some benchmarks that could grow or shrink with the number of PUs and work in different situations.
The second criterion is whether or not the workload is \textbf{representative}. As stated in the challenges, a workload must represent the production enviromenet or the experiment would be inconsistent~\cite{stephen_evaluate_2012}.

To meet these criteria, we looked at the "state of the art" and found the most common benchmarks used to test the performance of hardware. We then chose the ones that could be scaled up. 
Our candidate is  emph{NAS Parallel Benchmark} (NPB v3.3.1)~\cite{Bailey:1991:NPB:125826.125925}: one of the most used benchmarks for \emph{HPC}. We used the applications (\textsf{LU}), the \emph{Conjugate Gradient} (\textsf{CG}) and \emph{Embarrassingly Parallel} (\textsf{EP}) computation-intensive benchmarks in our experiments, with the \textsf{C} data class.
Further more we have used other applications to validate our results using more applications such as \texttt{Stress-ng v0.10.0},\footnote{\url{https://kernel.ubuntu.com/~cking/stress-ng}} \texttt{pbzip2 v1.1.9},\footnote{\url{https://launchpad.net/pbzip2/}} \texttt{linpack}\footnote{\url{http://www.netlib.org/linpack}} and \texttt{sha256 v8.26}\footnote{\url{https://linux.die.net/man/1/sha256sum}}.

\subsubsection{Metrics \& Measurement Tools}
Our metric for the accuracy of the test is the Standard deviation aka \textbf{STD} of the energy consumption. Therefore whether the tests consumes more or less energy is out of our scope.
To study this variation we need first a tool to measure the energy consumption. For this we used \textsc{PowerAPI}~\cite{DBLP:journals/jss/ColmantRKSFS18}, which is a power monitoring toolkit that is based on \emph{Intel Running Average Power Limit} (RAPL) \cite{Khan:2018:RAE:3199681.3177754}. The advantage of PowerAPI is that it reports the Energy consumption of CPU and DRAM at a socket level.

Our testbeds are run with a minimal version of Debian \,9 (4.9.0 kernel version)\footnote{\url{https://github.com/grid5000/environments-recipes/blob/master/debian9-x64-min.yaml}} where we install Docker (version 18.09.5), which will be used to run the RAPL sensor and the benchmark itself.
The energy sensor collects RAPL reports and stores them in a remote \textsc{MongoDB} instance, allowing us to perform \emph{post-mortem} analysis in a dedicated environment.
Using Docker makes the deployment process easier on the one hand, and provides us with a built-in control group encapsulation of the conducted tests on the other hand.
This allows \textsc{PowerAPI} to measures all the running containers, even the RAPL sensor consumption, as it is isolated in a container.

Every experiment is conducted on $100$ iterations, on multiple nodes and using the 3 NPB benchmarks we mentioned, with a warmup phase of 10 iterations for each experiment.
In most cases, we were seeking to evaluate the \emph{STandard Deviation} (STD), which is the most representative factor of the energy variation.
We tried to be very careful, while running our experiments, not to fall in the most common benchmarking "crimes"~\cite{DBLP:journals/corr/abs-1801-02381}.
As we study the STD difference of measurements we observed from empirical experiments, we use the bootstrap method~\cite{efron2000bootstrap} to randomly build multiple subsets of data from the original dataset, and we draw the STD density of those sets, as illustrated in Figure~\ref{fig:docker}.
Given the space constraints, this paper reports on aggregated results for nodes, benchmarks and workloads, but the raw data we collected remains available through the public repository we published.\footnote{\url{https://github.com/anonymous-data/Energy-Variation}}
We believe this can help to achieve better and more reliable comparisons.
We mainly consider 3 different workloads in our experiments: \textsf{single process}, 50\,\%, and 100\,\%, to cover the low, medium and high CPU usage when analyzing the studied parameters effect, respectively.
These workloads reflect the ratio of used PU count to the total available PU.

>>>>

\newpage
% To satisfy those three requirements. we elaborated a benchmarking protocol that aims to help reaserchers and practicioners in their experementations.
% for each aspect we present a set of guidelines and tools.
%% This is a conclusion maaan  !! 
\section{Improving accuracy by taming the energy variation}\label{sec:taming-the-energy-variation}

While the previous section aimed to ensure the reproducibility of our software energy consumption experiments, this section provides a collection of tips and tools to help increase the accuracy of these experiments.
We are well aware of the effect that hardware has on energy fluctuations.
However, we feel that there is still an opportunity for practitioners to minimize this energy variance by employing solely tuneable factors.
To that end, we conducted a series of empirical studies utilizing state-of-the-art recommendations to discover which controllable elements can limit the variations of benchmark energy usage.
% We will first start with evaluating the parameters mentioned by \cite{heinrich_predicting}.

In this section, we investigate the following research questions:
\begin{description}
    \item[\textsc{RQ}~1:] Does the benchmarking protocol affect the energy variation?
    \item[\textsc{RQ}~2:] How important is the impact of the processor features on the energy variation?
    \item[\textsc{RQ}~3:] What effect does the operating system have on energy variation?
    \item[\textsc{RQ}~4:] Does the choice of processor make a difference in reducing the energy variation?
\end{description}

@@

We used the platform Grid5000 (G5K)~\cite{grid5000,margery:hal-00965708}, which is a large, flexible testbed for experiment-driven research that is spread out across all of France, to meet these needs. Grid5000 has several clusters that are made up of 4 to 124 identical machines with different configurations for each cluster. For our experiment, we looked at four groups. Our main criterion was the CPU configuration.
The table~\ref{table:g5k}  below, presents a description of the four clusters that have been chosen for our experiments.

@@

Figure~\ref{fig:topo}  shows the detailed topology of a node in the cluster \textsf{Dahu} as an example.

@@

First, \textbf{scalability}: We wanted to learn as much as possible about the experiment in the time we had, so we needed some benchmarks that could grow or shrink with the number of PUs and work in different situations.
The second criterion is whether or not the workload is \textbf{representative}. As stated in the challenges, a workload must represent the production environment or the experiment would be inconsistent~\cite{stephen_evaluate_2012}.

To meet these criteria, we looked at the "state of the art" and found the most common benchmarks used to test the performance of hardware. We then chose the ones that could be scaled up. 
Our candidate is  emph{NAS Parallel Benchmark} (NPB v3.3.1)~\cite{Bailey:1991:NPB:125826.125925}: one of the most used benchmarks for \emph{HPC}. We used the applications (\textsf{LU}), the \emph{Conjugate Gradient} (\textsf{CG}) and \emph{Embarrassingly Parallel} (\textsf{EP}) computation-intensive benchmarks in our experiments, with the \textsf{C} data class.
Furthermore we have used other applications to validate our results using more applications such as \texttt{Stress-ng v0.10.0},\footnote{\url{https://kernel.ubuntu.com/~cking/stress-ng}} \texttt{pbzip2 v1.1.9},\footnote{\url{https://launchpad.net/pbzip2/}} \texttt{linpack}\footnote{\url{http://www.netlib.org/linpack}} and \texttt{sha256 v8.26}\footnote{\url{https://linux.die.net/man/1/sha256sum}}.

\subsubsection{Metrics \& Measurement Tools}
Our metric for the accuracy of the test is the Standard deviation aka \textbf{STD} of the energy consumption. Therefore whether the tests consume more or less energy is out of our scope.
To study this variation we need first a tool to measure the energy consumption. For this we used \textsc{PowerAPI}~\cite{DBLP:journals/jss/ColmantRKSFS18}, which is a power monitoring toolkit that is based on \emph{Intel Running Average Power Limit} (RAPL) \cite{Khan:2018:RAE:3199681.3177754}. The advantage of PowerAPI is that it reports the Energy consumption of CPU and DRAM at a socket level.

Our testbeds are run with a minimal version of Debian \,9 (4.9.0 kernel version)\footnote{\url{https://github.com/grid5000/environments-recipes/blob/master/debian9-x64-min.yaml}} where we install Docker (version 18.09.5), which will be used to run the RAPL sensor and the benchmark itself.
The energy sensor collects RAPL reports and stores them in a remote \textsc{MongoDB} instance, allowing us to perform \emph{post-mortem} analysis in a dedicated environment.
Using Docker makes the deployment process easier on the one hand, and provides us with a built-in control group encapsulation of the conducted tests on the other hand.
This allows \textsc{PowerAPI} to measure all the running containers, even the RAPL sensor consumption, as it is isolated in a container.

Every experiment is conducted on $100$ iterations, on multiple nodes and using the 3 NPB benchmarks we mentioned, with a warmup phase of 10 iterations for each experiment.
In most cases, we were seeking to evaluate the \emph{STandard Deviation} (STD), which is the most representative factor of the energy variation.
We tried to be very careful, while running our experiments, not to fall in the most common benchmarking "crimes"~\cite{DBLP:journals/corr/abs-1801-02381}.
As we study the STD difference of measurements we observed from empirical experiments, we use the bootstrap method~\cite{efron2000bootstrap} to randomly build multiple subsets of data from the original dataset, and we draw the STD density of those sets, as illustrated in Figure~\ref{fig:docker}.
Given the space constraints, we report on aggregated results for nodes, benchmarks and workloads. However, the raw data we collected is available through the public repository we published.\footnote{\url{https://github.com/anonymous-data/Energy-Variation}}
We believe this can help to achieve better and more reliable comparisons.
We mainly consider 3 different workloads in our experiments: \textsf{single process}, 50\,\%, and 100\,\%, to cover the low, medium and high CPU usage when analyzing the studied parameters effect, respectively.
These workloads reflect the ratio of used PU count to the total available PU.
